---
title: "Dada2_vis"
author: "Rebecca Clement"
date: "3/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries that we need
```{r message=FALSE}
library(dada2)
library(ggplot2)
library(phyloseq)
library(Biostrings)
library(dplyr)
```
Read in all the files that we need
```{r}
#objects with 2 after are the second set of samples we sequenced. Seqtab files have columns as DNA seqs, rows as IDs. tax files should have DNA seqs as rows and kingdom phylum class ranks as columns. 
seqtab<-readRDS("~/Box/Clement_TermiteGut_16S_0108/Dada2_rds_files/outfiles/seqtab_final.rds")
seqtab2<-readRDS("~/Box/Clement_TermiteGut_16S_0108/Dada2_rds_files/outfiles/seqtab_final2.rds")
tax<-readRDS("~/Box/Clement_TermiteGut_16S_0108/Dada2_rds_files/outfiles/tax_final.rds") # Need to go back and fix this file. It is showing bad stuff.
tax2<-readRDS("~/Box/Clement_TermiteGut_16S_0108/Dada2_rds_files/outfiles/tax_final2.rds") # This one hasn't finished running yet. Caroline will upload to BOX when it does
tax_silva<-readRDS("~/Box/Clement_TermiteGut_16S_0108/Dada2_rds_files/outfiles/tax_silva.rds")
tax_silva2<-readRDS("~/Box/Clement_TermiteGut_16S_0108/Dada2_rds_files/outfiles/tax_silva2.rds")

# Read in metadata file
term_meta<-read.csv("~/Box/Clement_TermiteGut_16S_0108/16S_metadata.csv")
```
## 1) Subset our metadata files so that it's equal with the seqtab files.
```{r}
samples1.out<-rownames(seqtab)
samples1_meta<-term_meta[term_meta$Termite_ID %in% samples1.out,]
rownames(samples1_meta)<-samples1.out[1:56]

samples2.out<-rownames(seqtab2)
samples2_meta<-term_meta[term_meta$Termite_ID %in% samples2.out, ]
rownames(samples2_meta)<-samples2.out[1:72]
```

## 2) Make phyloseq objects from seqtab, metadata and tax. Save these objects
```{r}
ps1<-phyloseq(otu_table(seqtab,taxa_are_rows = F),sample_data(samples1_meta),tax_table(tax_silva))

ps2<-phyloseq(otu_table(seqtab2,taxa_are_rows = F),sample_data(samples2_meta),tax_table(tax_silva2)) # Can't run this yet. Wait until Caroline adds thing.

#Store DNA names as dna to make more readable
dna <- Biostrings::DNAStringSet(taxa_names(ps1))
names(dna) <- taxa_names(ps1)
ps1 <- merge_phyloseq(ps1, dna)
taxa_names(ps1) <- paste0("ASV", seq(ntaxa(ps1)))
ps1

dna2 <- Biostrings::DNAStringSet(taxa_names(ps2))
names(dna2) <- taxa_names(ps2)
ps2 <- merge_phyloseq(ps2, dna2)
taxa_names(ps2) <- paste0("ASV", seq(ntaxa(ps2)))
ps2

```

## 3) Compare alpha diversity between samples
```{r}
plot_richness(ps1, x="Species", measures=c("Shannon", "Simpson"), color="Site")
plot_richness(ps2, x="Seq_ID", measures=c("Shannon", "Simpson"), color="Site")
```

## 4) Compare beta diversity with ordination. NMDS.
```{r}
# Transform data to proportions as appropriate for Bray-Curtis distances
ps1.prop <- transform_sample_counts(ps1, function(otu) otu/sum(otu))
ord.nmds.bray <- ordinate(ps1.prop, method="NMDS", distance="bray")
plot_ordination(ps1.prop, ord.nmds.bray, color="Species", title="Bray NMDS")

```
Same but for ps2
```{r}
ps2.prop <- transform_sample_counts(ps2, function(otu) otu/sum(otu))
ord.nmds.bray2 <- ordinate(ps2.prop, method="NMDS", distance="bray")
plot_ordination(ps2.prop, ord.nmds.bray2, color="Species", title="Bray NMDS")
```

## 5) Make a bar plot that shows the different samples.

```{r}
top20 <- names(sort(taxa_sums(ps1), decreasing=TRUE))[1:20]
ps1.top20 <- transform_sample_counts(ps1, function(OTU) OTU/sum(OTU))
ps1.top20 <- prune_taxa(top20, ps1.top20)
plot_bar(ps1.top20, x="Species", fill="Genus") + facet_wrap(~Species, scales="free_x")
```
Same but for ps2
```{r}
top20_2 <- names(sort(taxa_sums(ps2), decreasing=TRUE))[1:20]
ps2.top20 <- transform_sample_counts(ps2, function(OTU) OTU/sum(OTU))
ps2.top20 <- prune_taxa(top20, ps2.top20)
#plot_bar(ps2.top20, x="Species", fill="Genus") + facet_wrap(~Species, scales="free_x")
```

